### 2021-11-29 code/analysis_code/explorationscript.Rmd
Project part 5

---
title: "explorationscript"
author: "Gabriella Veytsel"
date: "November 12, 2021"
output: html_document
---
Load packages
```{r}
library(readxl) #for loading Excel files
library(here) #to set paths
library(tidyverse)
library(sf)#st_read to read shapefiles
library(tmap) #plot map
library(scales)
library(broom) #tidy
library(ggpubr) #add equation and r-square 
library(tidyr)
library(cowplot)
library(table1) #html table 1
```

Read RDS files from processingscript.R
```{r}
save_data_location1 <- here::here("data","processed_data","merge_shp_vac.rds")
save_data_location2 <- here::here("data","processed_data","cleandata_Trends.rds")
save_data_location3 <- here::here("data","processed_data","cleandata_Cases.and.Fatalities.rds")
save_data_location4 <- here::here("data","processed_data","cleandata_Vaccination.rds")
save_data_location5 <- here::here("data","processed_data","NewCasesTime_long.rds")
save_data_location6 <- here::here("data","processed_data","cleandata_Demographics.rds")

merge_shp_vac <- readRDS(save_data_location1)
cleandata_Trends <- readRDS(save_data_location2)
cleandata_Cases.and.Fatalities <- readRDS(save_data_location3)
cleandata_Vaccination <- readRDS(save_data_location4)
NewCasesTime_long <- readRDS(save_data_location5)
cleandata_Demographics <- readRDS(save_data_location6)
```

Epidemic curve
```{r}
epicurve <- ggplot(cleandata_Trends, aes(x=Date, y = `New Confirmed Cases`)) +
  geom_bar(stat = "identity", fill="lightblue3") +
  scale_x_date(date_breaks = "months" , 
               date_labels = "%b-%y", 
               limits = c(min(cleandata_Trends$Date), max = max(cleandata_Trends$Date))) + 
  theme(axis.text.x = element_text(angle = 45))

epicurve

#Save figure
figure_file_epicurve = here("results","epicurve.png")
ggsave(filename = figure_file_epicurve, plot=epicurve, height = 5, width = 8) 
```

Confirmed Case, by County
```{r}
tail(sort(merge_shp_vac$`Confirmed Cases`),10) #top 10 counties with the most confirmed cases
map_cases <- ggplot(merge_shp_vac) +
  geom_sf(aes(fill = `Confirmed Cases`)) +
  ggtitle("Confirmed Cases by County, as of 10/3/2021") +
  geom_sf_text(data = subset(merge_shp_vac, `Confirmed Cases` >= 67896), aes(label = COUNTY), size = 2.5) +
  scale_fill_gradient(low = "light blue", high = "red", name = "Confirmed Cases", labels = comma) 
#labels = comma removes scientific notation (package 'scales')

map_cases
#As expected, these are all the major cities. Instead, look at rates to compare counties with different population sizes

#Save figure
figure_file_map_cases = here("results","map_cases.png")
ggsave(filename = figure_file_map_cases, plot=map_cases, height = 12, width = 15) 
```

#Cumulative Confirmed Case Rate/100 Persons, by County
```{r}
tail(sort(merge_shp_vac$CumCaseRatePer100),10) #top 10 counties with the highest cumulative case rate

sites <- data.frame(latitude = c(35.4044, 34.5258, 34.1065, 33.6754, 29.8688,
                                  28.4117, 29.3343, 28.7410, 28.9089, 29.7525),
                    longitude = c(-101.9805, -100.1703, -101.8892, -101.7980, -101.1617,
                                  -99.8125, -99.8125, -100.3498, -97.8722, -94.6864))
#Potter: 35.4044° N, 101.9805° W
#Childress: 34.5258° N, 100.1703° W
#Hale: 34.1065° N, 101.8892° W
#Lubbock: 33.6754° N, 101.7980° W
#Val Verde: 29.8688° N, 101.1617° W
#Dimmit: 28.4117° N, 99.8125° W
#Uvalde: 29.3343° N, 99.8125° W
#Maverick: 28.7410° N, 100.3498° W
#Karnes: 28.9089° N, 97.8722° W
#Chambers: 29.7525° N, 94.6864° W

map_rates <- ggplot(merge_shp_vac) +
  geom_sf(aes(fill = CumCaseRatePer100), colour = "white", size = 0.05) +
  geom_point(data = sites, aes(x = longitude, y = latitude), size = 2, shape = 23, fill = "black") +
  geom_sf_text(data = subset(merge_shp_vac, CumCaseRatePer100 >= 19.1), aes(label = COUNTY), size = 4.5, nudge_y = -0.13) +
  ggtitle("Cumulative Confirmed Case Rate/100 Persons, by County, as of 10/3/2021") +
  scale_fill_gradient(low = "light blue", high = "red", name = "Cumulative Confirmed Case Rate/100 Persons", labels = comma)

map_rates

#Save figure
figure_file_map_rates = here("results","map_rates.png")
ggsave(filename = figure_file_map_rates, plot=map_rates, height = 12, width = 15) 
```

Epidemic Curve, by County
```{r}
#Too many counties for facet_wrap, narrow down to top 10 counties with highest case burden
NewCasesTime_long %>%
  group_by(County) %>%
  summarize(Sum = sum(NewCases)) %>%
  arrange(desc(Sum))

NewCasesTime_long_subset <- NewCasesTime_long %>%
  filter(County %in% c("Dimmit", "Hale", "Uvalde", "Maverick", "Karnes", "Childress", "Lubbock", "Val Verde", "Potter", "Chambers"))

epicurve_county <- ggplot(NewCasesTime_long_subset, aes(x=Date, y = NewCases)) +
  geom_bar(stat = "identity", fill="darkblue") +
  scale_x_date(date_breaks = "months" , 
               date_labels = "%b-%y", 
               limits = c(min(NewCasesTime_long$Date), max = max(NewCasesTime_long$Date))) + 
  scale_y_continuous(limits = c(min=0, max = 750)) +
  facet_wrap(~County, nrow = 2) + #2 rows 
  theme(axis.text.x = element_text(angle=50, hjust=1)) +
  labs(title = "Epidemic Curve", 
       subtitle = "Subset to counties with the highest case burden",
       y = "Cumulative Case Rate", x = "")

epicurve_county

#Save figure
figure_file_epicurve_county = here("results","epicurve_county.png")
ggsave(filename = figure_file_epicurve_county, plot=epicurve_county, height = 5, width = 15) 
```

```{r}
merge_shp_vac %>%
  dplyr::select(COUNTY, CumCaseRatePer100) %>%
  arrange(desc(CumCaseRatePer100))

merge_shp_vac_subset <- merge_shp_vac %>%
  dplyr::select(COUNTY, `Confirmed Cases`, Total, CumCaseRatePer100, Population_Density, Percent65Plus, PercentVaccinated_1dose, PercentVaccinated_full, PercentHispanic, PercentAsian, PercentBlack, PercentMale, AREA_SQKM, PercentMedCondition) %>%
  rename(Population = Total)
summary(merge_shp_vac_subset)

merge_shp_vac_df <- as.data.frame(merge_shp_vac_subset) %>%
  dplyr::select(-c(COUNTY, geometry))

merge_shp_vac_df_c <- as.data.frame(merge_shp_vac_subset) %>%
  dplyr::select(-c(geometry)) #has to be a dataframe without geometry for table 1

#Save dataset
save_data_location10 <- here::here("data","processed_data","merge_shp_vac_df_c.rds")
saveRDS(merge_shp_vac_df_c, file = save_data_location10)

table1 <- table1(~ . , data = merge_shp_vac_df, topclass="Rtable1-zebra") 
table1

#Save table
save_data_location7 <- here::here("results","summarytable.rds")
saveRDS(table1, file = save_data_location7)
```

#County characteristics, subset to counties with highest cumulative case rates
```{r}
table_county <- merge_shp_vac_df_c %>%
  filter(COUNTY %in% c("Dimmit", "Hale", "Uvalde", "Maverick", "Karnes", "Childress", "Lubbock", "Val Verde", "Potter", "Chambers" )) %>%
  group_by(COUNTY) %>%
  arrange(desc(CumCaseRatePer100))

#Save table
save_data_location6 <- here::here("results","table_county.rds")
saveRDS(table_county, file = save_data_location6)

library(vtable)
summarytable <- sumtable(merge_shp_vac_df_c, summ = c('mean(x)', 'min(x)', 'max(x)'),
         title="", out = "return")

#Save table
save_data_location17 <- here::here("results","summarytable.rds")
saveRDS(summarytable, file = save_data_location17)
```

Association between county-level characteristics and # confirmed cases 
```{r}
fit1 <- lm(`Confirmed Cases` ~ PercentHispanic, data = merge_shp_vac_subset, offset = Population)
summary(fit1) #Adjusted R-Squared = 0.56

fit2 <- lm(`Confirmed Cases` ~ Percent65Plus, data = merge_shp_vac_subset, offset = Population)
summary(fit2) #Adjusted R-Squared = 0.56

fit3 <- lm(`Confirmed Cases` ~ PercentMale, data = merge_shp_vac_subset, offset = Population)
summary(fit3) #Adjusted R-Squared = 0.56

fit4 <- lm(`Confirmed Cases` ~ PercentMedCondition, data = merge_shp_vac_subset, offset = Population)
summary(fit4) #Adjusted R-Squared = 0.56

fit5 <- lm(`Confirmed Cases` ~ PercentAsian, data = merge_shp_vac_subset, offset = Population)
summary(fit5) #Adjusted R-Squared = 0.57

fit6 <- lm(`Confirmed Cases` ~ PercentBlack, data = merge_shp_vac_subset, offset = Population)
summary(fit6) #Adjusted R-Squared = 0.56

fit7 <- lm(`Confirmed Cases` ~ AREA_SQKM, data = merge_shp_vac_subset, offset = Population)
summary(fit7) #Adjusted R-Squared = 0.56
#Why are they all about the same R-squared? I shouldn't use correlation tests (spearman, pearson) here because I need the offset

#Cumulative Case Rate
fit_rate <-  lm(sqrt(CumCaseRatePer100) ~ AREA_SQKM, data = merge_shp_vac_subset)
summary(fit_rate) #Why is the R-squared so low when I try this (with or without transformation)?

cor.test(merge_shp_vac_subset$CumCaseRatePer100, merge_shp_vac_subset$AREA_SQKM, method = "pearson")

#I am so confused...
#Cumulative Case Rate
ggplot(merge_shp_vac_subset, aes(y = log(CumCaseRatePer100), x = AREA_SQKM)) +
       geom_point() +
       geom_smooth(method = "lm")

#Confirmed Cases
ggplot(merge_shp_vac_subset, aes(y = log(`Confirmed Cases`), x = AREA_SQKM)) +
       geom_point() +
       geom_smooth()
hist(merge_shp_vac_subset$AREA_SQKM) #is there just not enough range?
#When I don't add lm, looks like a polynomial, but fit was worse when added ^2 or ^2 and ^3

#Initially, I fit confirmed case against population numbers (# 65+) instead of percentages ((%65+) and the R-squared was like 0.99, so I think this is confounding because it's just getting at population size

#Confirmed Case
hist(log(merge_shp_vac_subset$`Confirmed Cases`)) #LOG TRANSFORMATION GIVES THE OUTCOME A NORMAL DISTRIBUTION! Since sample size is pretty small (254), law of large numbers doesn't apply.
shapiro.test(log(merge_shp_vac_subset$`Confirmed Cases`))
#Did try fitting log(Cofirmed Cases), but r-squared and AIC was worse

#Cumulative Case Rate
hist(merge_shp_vac_subset$CumCaseRatePer100)  
shapiro.test(sqrt(merge_shp_vac_subset$CumCaseRatePer100)) #SQUARE ROOT TRANSFORMATION GIVES THE OUTCOME A NORMAL DISTRIBUTION
```

```{r}
# p1 <- ggscatter(merge_shp_vac_subset, x="PercentHispanic", y="CumCaseRatePer100",
#           add = "reg.line", conf.int = TRUE,
#           cor.coef = TRUE, cor.method = "spearman")
#
# p2 <- ggscatter(merge_shp_vac_subset, x="Percent65Plus", y="CumCaseRatePer100",
#           add = "reg.line", conf.int = TRUE,
#           cor.coef = TRUE, cor.method = "spearman")
# p3 <- ggscatter(merge_shp_vac_subset, x="PercentMale", y="CumCaseRatePer100",
#           add = "reg.line", conf.int = TRUE,
#           cor.coef = TRUE, cor.method = "spearman")
# p4 <- ggscatter(merge_shp_vac_subset, x="PercentBlack", y="CumCaseRatePer100",
#           add = "reg.line", conf.int = TRUE,
#           cor.coef = TRUE, cor.method = "spearman")
# p5 <- ggscatter(merge_shp_vac_subset, x="PercentAsian", y="CumCaseRatePer100",
#           add = "reg.line", conf.int = TRUE,
#           cor.coef = TRUE, cor.method = "spearman")
# p6 <- ggscatter(merge_shp_vac_subset, x="PercentMedCondition", y="CumCaseRatePer100",
#           add = "reg.line", conf.int = TRUE,
#           cor.coef = TRUE, cor.method = "spearman")
# p7 <- ggscatter(merge_shp_vac_subset, x="AREA_SQKM", y="CumCaseRatePer100",
#           add = "reg.line", conf.int = TRUE,
#           cor.coef = TRUE, cor.method = "spearman")
# 
# p_combined <- plot_grid(p1, p2, p3, p4, p5, p6, p7) 
```